A Study of Immersion Factors Using an Augmented Reality Simulation

### 2009-06-24 trunk/Paper/paper.tex
more figures

In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.  Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.  It also gives us the ability to remove unnecessary factors from generating noise in our data: illumination

By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.  The simulation is a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.  We wish to answer the following questions.  What effect does varying the augmented FOV have on user tracking performance?  What effect do different dropout lengths have on performance?  What are the interactions between these two variables?

@@

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
%	\caption{Interaction plot of the effect of field of view and registration dropouts on the number of times the participant becomes lost.}
%\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
\end{figure}

@@

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
%\end{figure}

@@

We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.  Table \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view had a very significant effect on performance using this metric.  
% TODO: interaction?

% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.

Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.  Performance decreases between zero and one seconds, but seems to level off between one and two.  We ran a post-hoc analysis to determine which conditions were significantly different.  Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).  This suggests that we may be seeing a thresholding effect above one second sensor dropouts.

This suggests that we have reached a threshold at one second.  We had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.  Figure 8 shows the data.  There wasn't a significant difference between the zero length and half second length conditions.  For future work, we could find the threshold between 0.5 and 1.

@@

\begin{figure}[htb]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
Metric & Factor & F value & Pr \\
\hline
Time Tracking & Dropout Length & 20.41 & $<0.001$\\
Time Tracking & Field of View & 13.549 & $<0.001$\\
Number of Times Lost & Dropout Length & 61.816 & $<0.001$\\
\hline
\end{tabular}
 \caption{\label{fig:anova}Factors which have a significant effect.}
\end{figure}

@@

\section*{Acknowledgements}
Cha Lee, Steffan Gauglitz, funding, and all of our participants.

>>>>

In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.  Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.  It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.

The AR simulation is built as a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.  We wish to answer the following questions.  What effect does varying the augmented FOV have on user tracking performance?  What effect do different dropout lengths have on performance?  What are the interactions between these two variables?

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=3.5in]{figures/tt_deadlen.pdf}
	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
\end{figure}
% TODO: make the units seconds and change the labels on the bottom of this figure

\begin{figure}[t]
	\centering
	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
\end{figure}

@@

@@

We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.  Figure \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view had a very significant effect on performance using this metric.  
% TODO: interaction?

We also used a two-way {ANOVA} to examine the number of times the participant loses the target.
% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
\begin{figure}[h]

@@

\centering
\begin{tabular}{|c|c|c|c|}
\hline
%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
Metric & Factor & F value & Pr \\
\hline
Time Tracking & Failure Length & 28.904 & $<0.001$\\
Time Tracking & Field of View & 15.800 & $<0.001$\\
Number of Times Lost & Failure Length & 8.5008 & $<0.001$\\
Number of Times Lost & Field of View & 13.060 & $<0.001$\\
\hline
\end{tabular}
 \caption{\label{fig:anova}Factors which have a significant effect.}
\end{figure}

Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.  Performance decreases between zero and one seconds, but seems to level off between one and two.  We ran a post-hoc analysis to determine which conditions were significantly different.  Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p=??), but not significantly different from each other (p=??).  This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
% TODO: table with values

To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.  Figure \ref{fig:extra} shows a box plot of time tracking versus sensor dropout length for these seven participants.  Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.  
%TODO: table with values
With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
%For future work, we could find the threshold between 0.5 and 1.

\begin{figure}[t]
	\centering
	\includegraphics[width=3.5in]{figures/extra.pdf}
	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.  We collected this data from seven participants, using the medium field of view.}
\end{figure}

@@

\section*{Acknowledgements}
Cha Lee, Steffen Gauglitz, funding, and all of our participants.
